---
title: "Public Datasets"
author: "Marie"
date: "2018-05-30"
slug: "public-datasets"
categories: [My-Learning-Path]
tags: []
draft: false
banner: "img/banners/online_data.jpg"
---



**[Titanic](https://www.coursera.org/learn/machine-learning)** 



**[Ames Housing](https://ww2.amstat.org/publications/jse/v19n3/decock.pdf)**

  * **SampleSource**: 2,930
  * **Features**: 81
  * **Response Variable**: SalePrice
  * **Response Type**: Continous
  * **Comments**:
      * See also [https://www.kaggle.com/c/house-prices-advanced-regression-techniques](https://www.kaggle.com/c/house-prices-advanced-regression-techniques)

```{r load_dataset_ames, include=FALSE}
ames <- readr::read_tsv("../../static/datasets/amesHousing/AmesHousing.txt")
str(ames)
```

**[Music Genre](http://tunedit.org/challenge/music-retrieval/genres)**

  + **Objective**: Classify music into 6 categories
  + **Sample size**: 12,495
  + **Features**: 191 
  + **Response Type**: Categorical
  + **Comments**: 
    * Unbalanced Response
    * Correlated predictors 
    * Predictors spanned different scales of measurement
    * Samples not independent (15-20 pieces of music for each performer, then each piece split into 20 segments)
    * Closed Contest, : 0.87507 standard accuracy (correctly predictions/ n).  Baseline provided: generated using 1-NN algorithm without any preprocessing


```{r load_dataset_genres, include=FALSE}
genres <- readr::read_csv("../../static/datasets/musicGenres/genresTrain.csv")
```

**[Grant Applications](https://www.kaggle.com/c/unimelb/data)**

  + **Objective**: Predict probaiblity of success of a grant application
  + **Sample size**: 8,707
  + **Features**: 249
  + **Response Type**: Categorical
  + **Comments**:
    * Missing values (83%)
    * Dependent samples; same grant writters occurred mlutiple times

```{r load_dataset_grants, include=FALSE}
require(tidyverse)
grants <- readr::read_csv("../../static/datasets/grantApplications/unimelb_training.csv") %>%
  select(-X252)
```


**[Hepatic Injury](http://appliedpredictivemodeling.com/data/)**

  + **Objective**:
  + **Sample Size**: 281
  + **Features**: 376 (184 bio, 192 chem)
  + **Response Type**: Categorical
  + **Comments**:
     * Data available within R package `AppliedPredictiveModeling` as `bio` and `chem`.
     * Unbalanced response

```{r load_dataset_hepatic, include=FALSE}
require(AppliedPredictiveModeling)
data(bio)
data(chem)
```


**[Permeability](http://appliedpredictivemodeling.com/data/)**

  + **Sample Size**: 165
  + **Features**: 1,107
  + **Response Type**: Continuous
  + **Comments**: 
    * Data available within R package `AppliedPredictiveModeling` as `permeability`
    * Correlated samples
    * Sparse data set
```{r load_dataset_permeability, include=FALSE}
require(AppliedPredictiveModeling)
data(permeability)
#see also fingerprints
```

**[Chemical Manufacturing Process](http://appliedpredictivemodeling.com/data/)**

  + **Sample size**: 177
  + **Features**: 57
  + **Response Type**: Continuous
  + **Comments**
    * Data available within R package `AppliedPredictiveModeling` as `ChemicalManufacturingProcess`
    * Correlated data
    * Missing Values
  
```{r load_dataset_chemicalmanufacturingprocess, include=FALSE}
require(AppliedPredictiveModeling)
data(ChemicalManufacturingProcess)
```


